---
title: "Analyzing Career Mobility in the CMap Dataset: Practical Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analyzing Career Mobility in the CMap Dataset: Practical Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(cmapr)
library(tidyverse)

# Set dataset_dir to where you downloaded the CMap data
# Change this path to match your local system!
dataset_dir <- "~/cmap_data/dataset"
```

# **Analyzing Career Mobility in the CMap Dataset: Practical Examples**

This vignette provides hands-on examples for analyzing job specialization, title mapping, and promotion networks using the **cmapr** package and the [CMap Career Mobility Dataset](https://www.nature.com/articles/s41597-025-05526-3).

The examples are designed to help you efficiently explore and visualize the data, leveraging tidyverse tools for clarity and reproducibility.

------------------------------------------------------------------------

## **1. Loading and Inspecting Specialization Data**

The specialization index (SI) quantifies how distinctive a job title is within its sector.

```{r}
si_data <- load_sector_specialization(file.path(dataset_dir, "titles/si"))
si_data %>%
  group_by(sector) %>%
  arrange(desc(si)) %>%
  slice_head(n = 10) %>%
  select(sector, title, si)
```

**Tip:**\
To find the most specialized roles in a sector, filter by high SI scores.

```{r}
si_data %>%
  filter(sector == "Information Technology") %>%
  arrange(desc(si)) %>%
  slice_head(n = 5)
```

------------------------------------------------------------------------

## **2. Mapping and Summarizing Job Titles**

The title mapping data helps you understand how raw job titles are standardized and grouped.

```{r}
title_map <- load_title_map(file.path(dataset_dir, "titles/map"))
title_map %>%
  count(sector, title_simplified, sort = TRUE)
```

**Tip:**\
Use this mapping to collapse variant job titles and compute sector-level frequencies.

```{r}
title_map %>%
  group_by(sector, title_simplified) %>%
  summarise(total = sum(frequency_simplified, na.rm = TRUE)) %>%
  arrange(sector, desc(total))
```

------------------------------------------------------------------------

## **3. Exploring Promotion Networks**

Promotion networks show observed transitions between job titles, both validated (manually reviewed) and unvalidated (algorithmic).

### **Validated Promotions**

```{r}
validated_edges <- load_validated_promotions("edges", file.path(dataset_dir, "promotions/validated"))
validated_nodes <- load_validated_promotions("nodes", file.path(dataset_dir, "promotions/validated"))
```

### **Unvalidated Promotions**

```{r}
unvalidated_edges <- load_unvalidated_promotions("edges", file.path(dataset_dir, "promotions/unvalidated"))
unvalidated_nodes <- load_unvalidated_promotions("nodes", file.path(dataset_dir, "promotions/unvalidated"))
```

**Tip:**\
Investigate the most frequent promotion paths in a sector:

```{r}
unvalidated_edges %>%
  filter(sector == "Healthcare") %>%
  arrange(desc(frequency)) %>%
  slice_head(n = 5) %>%
  select(from, to, frequency, promotion_prob)
```

------------------------------------------------------------------------

## **4. Visualizing Career Transitions**

You can visualize promotion networks interactively (if HTML files are available):

```{r, eval = FALSE}
load_validated_promotions("network", file.path(dataset_dir, "promotions/validated"), open_html = "US_finance.html")
```

Or plot a simple summary using ggplot2:

```{r}
top_edges <- unvalidated_edges %>%
  group_by(sector) %>%
  arrange(desc(frequency)) %>%
  slice_head(n = 5)

ggplot(top_edges, aes(x = reorder(from, frequency), y = frequency, fill = to)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sector, scales = "free") +
  labs(title = "Most Frequent Career Transitions by Sector",
       x = "From Title",
       y = "Frequency") +
  theme_minimal(base_size = 12)
```

------------------------------------------------------------------------

## **5. Further Exploration and Research**

-   Combine `cmapr` data tables with external labor market data for advanced modeling.

-   Use the specialization index to identify emerging or declining roles.

-   Explore cross-sector mobility patterns to inform workforce development.

**References:**

-   [CMap Career Mobility Dataset - Nature Scientific Data (2025)](https://www.nature.com/articles/s41597-025-05526-3)

-   [cmapr package documentation](https://github.com/farach/cmapr)

------------------------------------------------------------------------

These examples provide a foundation for your own research and exploration of global career mobility.\
For more advanced analyses, see additional articles or the package reference.
