---
title: "Get Started with cmapr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get Started with cmapr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(cmapr)
library(dplyr)
```

# **Introduction**

**cmapr** is a modern R package for tidy, reproducible analysis of the [CMap Career Mobility Dataset](https://zenodo.org/records/15260189)—the largest open research dataset of global career transitions, created for the study ["A global career mobility map from 170 million job transitions"](https://www.nature.com/articles/s41597-025-05526-3) published in *Scientific Data* (Nature).

The CMap dataset provides anonymized, aggregated records of **170 million real-world job transitions** across countries, sectors, and job titles. It includes standardized mappings of job titles, sector and regional labels, specialization indices, and directed promotion networks, enabling powerful labor market research and modeling.

**cmapr** makes it easy to access, manipulate, and analyze this data using the tidyverse ecosystem.

------------------------------------------------------------------------

## **Installation**

Install from GitHub using `remotes`:

```{r, eval=FALSE}
install.packages("remotes")
remotes::install_github("farach/cmapr")
```

------------------------------------------------------------------------

## **Downloading the Data**

The full CMap dataset (\~130MB) is **not bundled** with cmapr due to size and licensing. Instead, you can download the official release directly from Zenodo.

**Tip:**\
Choose a permanent folder for the data so you only need to download it once. For example, use `"~/cmap_data"` (your home directory), or a project-specific path.

```{r}
library(cmapr)
dataset_dir <- download_cmap_data("~/cmap_data")
```

-   The function will download and unzip the data.

-   If the data already exists, it will skip re-downloading by default.

-   To **force overwrite/download**, set `overwrite = TRUE` :

    ```{r}
    dataset_dir <- download_cmap_data("~/cmap_data", overwrite = TRUE)
    ```

-   Once downloaded, you can re-use this folder for all your analyses.

**Efficient Data Management:**

-   Store data in a location that’s backed up and easily accessible for all your R projects.

-   On shared or cloud systems, consider a workspace directory or mounted drive.

-   If you use RStudio projects, you may want to keep `"cmap_data"` outside your main package folder to avoid accidental deletion.

------------------------------------------------------------------------

## **Loading and Exploring the Data**

After downloading, load the data into tidy tibbles for analysis:

```{r}
result <- load_cmap_data(base_path = dataset_dir)
model_data <- result$model_data      # Main transitions and features
metadata   <- result$metadata        # Job title and sector metadata
```

You now have instant access to the core CMap tables, ready for use with dplyr, tidyr, and ggplot2.

------------------------------------------------------------------------

## **Next Steps**

See the next article for practical examples of data exploration with cmapr!
